# Conclusion
<<<<<<< ECON_RESULTS
In the above analysis we have explore the relation of internet availability and various socioeconomic features of neighborhoods throughout the five boroughs of New York City. In this analysis we have discovered some of concerning discrepancies with internet availability, interesting trends regarding the distribution of people and internet, and some statistics that seem encouraging regarding Mayor de Blasio's Internet Master Plan announced in 2020.

Regarding internet coverage around the city we can see that the best internet coverage is in Manhattan. Staten Island and Brooklyn have mid to high level internet adoption while Queens comes in slightly higher than the Bronx, but both at the bottom of the list. We came to this conclusion using data on internet adoption and availability, internet speed availability, public hot spot availability and mobile internet dependency. 

When looking at the distribution of of internet across New York City, one of the most concerning findings is the relation of mean income and internet availability and adoption. Referring to the graph above we can see that overall, a higher income in a neighborhood correlates with better internet availability and adoption. This signifies a continued socioeconomic imbalance of internet adoption in New York City. This trend is especially evident in The Bronx with the highest poverty rates, lowest income and the lowest internet adoptiona nd availability, contrasted with Manhattan, with the highest income levels as well as the best internet coverage and adoption. 
=======
Conclusion can start here .....

Below this is just for review on potential charts to keep

```{r}

library(forcats)
library(tidyr)
library(dplyr)
library(RColorBrewer)
library(ggpubr)
library(ggplot2)
library(ggridges)
library(GGally)
library(ggrepel)
library(vcd)
library(ggthemes)

```

```{r}
data = read.csv(file='data/clean/combined_dataset_no_weird.csv')

data_snap <- select(data, c('Borough',
                          'Neighborhood.Name',
                          'Total.Population',
                          'Total.Number.of.Households',
                          'Per.Capita.Income..in...',
                          'Home.Broadband.Adoption..', 
                          'Residential.Broadband.Choice.Average.by.NTA',
                          'Mobile.Dependent.Households..Percentage.of.Households.' ,
                          'Public.Wi.Fi.in.NYC.Count', 
                          'Mobile.Broadband.Adoption..',
                          'English.only.language..',
                          'Language.other.than.English..',
                          'Native.born.in.US..',
                          'Foreign.born..',
                          'Completed.Bachelors.degree.or.higher..',
                          'Median.House.gross.rent..in...',
                          'Median.House.value..in...', 
                          'Owner.occupied.Housing.unit..',
                          'Total.occupied.Housing.unit..',
                          'Renter.occupied.Housing.unit..',
                          'Work.from.Home...Capable',
                          'Work.from.Home...Incapable',
                          'plurality_racial_group'
                          ))
data_snap$Per.Capita.Income..in... <- data_snap$Per.Capita.Income..in.../1000
data_snap$Mobile.Dependent.Households..Percentage.of.Households. <- data_snap$Mobile.Dependent.Households..Percentage.of.Households. * 100
data_snap$Mobile.Broadband.Adoption.. <- data_snap$Mobile.Broadband.Adoption.. * 100
data_snap$Home.Broadband.Adoption.. <- data_snap$Home.Broadband.Adoption.. * 100

```

```{r}
# Internet Access Ratings around Rented Stays
iar <- ggplot(data, aes(fill=Borough, x=Residential.Broadband.Choice.Average.by.NTA, y=Renter.occupied.Housing.unit..)) +
    geom_line(aes(color=Borough), alpha=.5, size =8) +
    ggtitle("Internet Access Ratings around Rented Stays") +
    xlab("Region Broadband Rating by NTA") +
    ylab("Rented Housing Units in %") +
    theme_economist() + 
    scale_color_economist() +
    theme(axis.title = element_text(size = 5),
    axis.text = element_text(size = 5))

iar
```
Above plot analyses the broadband effectiveness in and around rented stays. Students, employed or unemployed people need to ensure the rent they pay give them the best facilities that's available in the region. As most of the people are immigrants, their internet connectivity concern may have a role in determining their quality of life as they may choose to live away from home.
We can observe the Brooklyn borough has wide variance in the internet connectivity spread and broadband rating in neighborhoods, however Manhattan on the other end has pretty much better access and connectivity due to more activity and internet dependent businesses that prevail in the area.The Staten island has an average NTA rating of 1.8 which is less than average of other boroughs like the Queens and Manhattan but far better than the Bronx area.


```{r}
# Borough wise Household Distribution
bhd <- ggplot(data, aes(x=Borough, y=Total.Number.of.Households)) +
   geom_boxplot() + facet_wrap(~ plurality_racial_group) +
   theme_economist() + 
   scale_color_economist() +
   theme(axis.title = element_text(size = 7),
        axis.text = element_text(size = 7))

bhd
```


```{r}
pcolors <- wsj_pal(palette = "colors6")
freq_count <- data %>%
  group_by(Borough,plurality_racial_group) %>%
  summarize(Freq = n())

# Basic Mosaic Plot of Racial groups residing in different boroughs
fig_mos <- vcd::mosaic( plurality_racial_group ~ Borough,data = freq_count,
             set_labels=list(plurality_racial_group = c("W_e","H/L","A_n","B/A_n"),
                             highlighting_fill = pcolors,
                             Borough = c("B","M","Q","SI","Bx")),direction = c("v","h"))

fig_mos

```

```{r}
# Enhanced Mosaic Plot of Racial groups residing in different boroughs
cex.axis = 0.5
adv_mos <- mosaicplot(data$Borough ~ data$plurality_racial_group, 
           main = "Mosaic Distribution of Racial Groups",
           sub = "",
           ylab = "Plurality Racial Groups",
           xlab = "Boroughs",
           las = 1,
           border = "chocolate",
           cex.axis = 0.5,
           shade = TRUE)

adv_mos
```

```{r}
# Democratized Internet Access Spread
dias <- ggplot(data, aes(Home.Broadband.Adoption..,Total.Number.of.Households)) +
  geom_point(aes(color = plurality_racial_group)) +
  theme_grey(10) +
  facet_grid(~Borough) +
  theme(strip.text = element_text(size = 10, color = "blue"),legend.position = "bottom") +
  xlab("Broadband Adoption %") +
  ylab("Number of Households in various neighborhoods") +
  ggtitle("Check for Democratized Internet Access Spread") +
  theme_wsj(base_size = 7,) +
  scale_colour_wsj("colors6") +
  theme(axis.title=element_text(size=7))

dias
```

```{r}
# Home Internet Users Lifestyle
wsj_plot <- ggplot(data, aes(fill=Borough, y=Home.Broadband.Adoption.., x=Median.House.gross.rent..in...)) +
    geom_line(aes(color=Borough), alpha=.5, size =2) +
    ggtitle("Home Internet Users Lifestyle") +
    ylab("Percentage of Home Broadband Users (%) ") +
    xlab("House Gross Rent (in $)") +
    theme_wsj(base_size = 10,) +
    scale_colour_wsj("colors6") +
    theme(axis.title=element_text(size=9))

wsj_plot

```

```{r}

# Broadband Ratings in HCOL & LCOL
wsj_plot2 <- ggplot(data, aes(fill=Borough, x=Residential.Broadband.Choice.Average.by.NTA, y=Median.House.value..in...)) +
    geom_point(aes(color=Borough), alpha=.5, size =7) +
    ggtitle("Broadband Ratings in HCOL & LCOL") +
    xlab("Broadband Rating by NTA ") +
    ylab("House Value (in $)") +
    theme_wsj(base_size = 9,) +
    scale_colour_wsj("colors6") +
    theme(axis.title=element_text(size=9))

wsj_plot2

```


```{r}

# Parallel Coordinates Plot of Variables
ppcoord_plot <- ggparcoord(data, columns = c(4,7,6,8), alphaLines = 0.5,scale = "uniminmax", splineFactor = 10, showPoints = TRUE) +
  theme(legend.position = "none",plot.title = element_text(hjust = 0.2)) +
  scale_color_manual(values = c("grey70")) +
  ggtitle("Parallel plot of different variables for Internet Households") +
  geom_vline(xintercept = 1:4, color = "black") +
  theme(axis.text.x = element_text(angle=10))

ppcoord_plot
```


