# Results

```{r, echo=FALSE, include=FALSE}
# libraries
library(dplyr)
library(leaflet)
library(sp)
library(ggmap)
library(maptools)
library(broom)
library(httr)
library(rgdal)
library(ggplot2)
library(tigris)
library(RColorBrewer)
library(ggpubr)
library(ggridges)
library(GGally)
library(RColorBrewer)
library(ggrepel)
library(forcats)
library(tidyr)
# import data
library(ggthemes)
data = read.csv(file='data/clean/combined_dataset.csv')
```
Home Internet Adoption Map:
```{r, echo=FALSE}
# import map data
nyc_neighborhoods <- readOGR('data/map/NTA_map.geojson')

current <- select(data, c('Borough', 'Neighborhood.Name', 'Home.Broadband.Adoption..'))
names(current)[1] <- 'borough'
names(current)[2] <- 'neighborhood'
names(current)[3] <- 'points'

current$borough_num <- unclass(current$borough)
current$opacity_lvl <- current$points
current$opacity_lvl <- current$opacity_lvl^3

map_data <- geo_join(nyc_neighborhoods, current, "ntaname", "neighborhood")

pal <- colorFactor(palette = 'Set1',
                    unique(map_data@data$boroname))

leaflet(map_data) %>%
  addTiles() %>%
  addPolygons(popup = paste("Neighborhood:", map_data@data$ntaname, "<br>",
                            "Borough:", map_data@data$boroname, "<br>",
                            "Internet Adoption Rate:", paste(map_data@data$points*100, '%', sep=''), "<br>"),
              fillColor = ~pal(map_data@data$boroname),
              fillOpacity = map_data@data$points^3 + .1,
              highlightOptions = highlightOptions(weight = 3),
                                  color = "Black",
                                  weight = 1) %>%
  addProviderTiles("CartoDB.Positron") %>%
  addLegend("bottomright", pal = pal, values = ~map_data@data$boroname, title="Boroughs")
```

```{r}
data <- read.csv("./data/clean/combined_dataset_no_weird.csv")
```

# Racial Stratification in Internet Infrastructure

It is impossible to discuss disparity in New York City without talking about race and national origin. New York has at various times been a key port in the trans-Atlantic slave trade, one of the largest destinations of the Great Migration, a "golden door" for myriad "huddled masses" sailing from the world's teeming shores, an origin of White Flight, and a battleground of gentrification. The present-day product of so much history is a complex geography of clustering, mixing, and segregation between races tracing their ancestry all over the world. This racial geography has informed decades and centuries of infrastructure construction and maintenance, including infrastructure to connect New Yorkers to the Internet. In 2020, a novel coronavirus rendered Internet connectivity more important than ever. In doing so, it also highlighted the degree of digital division in New York City's access to the Internet.

```{r}
data$plurality_racial_group <- as.factor(data$plurality_racial_group)
data$plurality_racial_group <- fct_relevel(
  data$plurality_racial_group,
  "White", "Asian", "Black or African American", "Hispanic or Latino"
)
broadband_vs_race <- ggplot(
  data, aes(x=100*Home.Broadband.Adoption.., y=fct_rev(plurality_racial_group))
) +
  geom_density_ridges(scale = 2) +
  theme_bw() +
  labs(
    title="Adoption of Home Broadband by Neighborhood Racial Plurality",
    x="% Households with Wired Broadband Internet",
    y="Plurality Racial Group"
  )
broadband_vs_race
```

Perhaps the most significant factor in one's ability to work or learn remotely is the quality of one's home Internet connection. There are many ways for a household to get connected, including wired options such as fiber-optic, coaxial cable, and dial-up, as well wireless options such as through a cellular network or via satellite. In general, wired options tend to be faster and more reliable. They are, however, more costly for the provider to install and for the end-user to subscribe to. Thus, one key metric of a neighborhood's ability to weather work-from-home and remote-learning is the portion of its households with a wired broadband Internet connection.

While at least half of households in almost every neighborhood in the city have access to broadband, there is a clear racial pattern to how close a neighborhood is to universal broadband. On average, over three quarters of households in neighborhoods where White-identifying residents made up the largest racial group had access to and made use of broadband Internet service in 2021. For neighborhoods of every other racial-plurality group, the average portion of households having adopted broadband Internet was lower, with neighborhoods home to a plurality Hispanic/Latino population having the lowest average broadband adoption rate of all.

```{r}
mobile_reliant_vs_race <- ggplot(data, aes(x=100*Mobile.Dependent.Households..Percentage.of.Households., y=fct_rev(plurality_racial_group))) +
  geom_density_ridges(scale = 2) +
  theme_bw() +
  labs(
    title="Mobile-Dependence by Neighborhood Racial Plurality",
    x="% Households Reliant on Wireless Mobile Internet",
    y="Plurality Racial Group"
  )
mobile_reliant_vs_race
```

An inverse measure of neighborhood Internet infrastructure is the proportion of households which rely on a wireless mobile connection. While advances in wireless technology have narrowed the speed gap between cellular-based Internet connections and their terrestrial counterparts, the former are still not able to handle multi-party video calls, HD streaming, and other data-intensive tasks as well as the latter. Mobile-dependent households thus enjoy diminished connectivity, especially when one considers that almost all households with a wired connection *also* have access to wireless mobile internet.

As can be expected, neighborhoods with a lower portion of households with a wired connection tend to have a higher portion of mobile-dependent households. The result is a similar racial disparity in both. In several plurality-Black neighborhoods, over a quarter of households are reliant on a wireless connection. No plurality-White neighborhood has over 15% household mobile-reliance. By all measures of central tendency, a smaller fraction of households in plurality-White neighborhoods rely on their cell phones for Internet than in any other racial-plurality neighborhoods.

Mayor de Blasio opened New York City's *Internet Master Plan*'s with a letter. In it, he lamented that "this digital divide, like so many other aspects of life in New York City, leaves a significant part of our population at a major disadvantage." One year after the letter's publishing, the *Plan*-directed survey of Internet access revealed that the city still has more of a gap to close the divide along all too familiar racial lines.

# 

# The Golden Door to Cyberspace: Internet Infrastructure of New York's Foreign-Born

New York City has always prided itself on its global connectivity. For over a century since the poet Emma Lazarus wrote of the city's beloved Lady Liberty "From her beacon-hand / Glows world-wide welcome", people from around the globe have made New York their home. Today's city population was born in over 150 countries. Like many people moving to unfamiliar places, many chose to live around those of similar backgrounds. As a consequence, 34% of the median City neighborhood, Bay Ridge, Brooklyn, was born outside of the United States. In over two dozen neighborhoods more than half of the population was born abroad. 10% of residents were born elsewhere even in the quiet suburb Rossville-Woodrow near Staten Island's southwestern tip, which has the second-lowest proportion of foreign-born residents of any City neighborhood. While immigrants make up a portion of all New York neighborhood populations, the make a more significant contribution to some neighborhood populations than to others.

```{r}
data3 <- data %>% rename(
  Pct.Foreign.Born = Foreign.born..,
  Home.Broadband.Adoption = Home.Broadband.Adoption..,
  Mobile.Dependent = Mobile.Dependent.Households..Percentage.of.Households.,
  Broadband.Providers = Residential.Broadband.Choice.Average.by.NTA,
  Public.Hotspots = Public.Wi.Fi.in.NYC.Count,
  Broadband.Providers = Residential.Broadband.Choice.Average.by.NTA,
  Mobile.Broadband = Mobile.Broadband.Adoption..
)

data3$foreign.quantile <- cut(
  data3$Pct.Foreign.Born,
  quantile(data3$Pct.Foreign.Born),
  include.lowest=TRUE,
  labels=FALSE,
  breaks=10,
)

par_coord_necessary = c(
  "foreign.quantile",
  "Pct.Foreign.Born",
  "Mobile.Dependent",
  "Home.Broadband.Adoption",
  "Broadband.Providers"
)
par_coord_cols = par_coord_necessary[
  !par_coord_necessary %in% c("foreign.quantile")
]
data3 <- data3 %>% select(all_of(par_coord_necessary))
data3
data3 <- data3 %>%
  group_by(foreign.quantile) %>%
  summarise_at(
    c(
      "Pct.Foreign.Born",
      "Mobile.Dependent",
      "Home.Broadband.Adoption",
      "Broadband.Providers"
  ),
  mean
)

foreign_born_par_coord <- ggparcoord(
  data3,
  columns=par_coord_cols,
  order=c(1, 2, 3, 4, 5, 6, 7),
  alphaLines = 0.7,
  scale = "center",
  splineFactor = 0,
  groupColumn = 2,
  mapping=aes(color=as.factor(foreign.quantile))
) +
  labs(
    title="Broadband Availability and Adoption in Foreign/Native-Born Neighborhoods",
    y="Scaled Value"
  ) +
  guides(color = guide_legend(title = "% Foreign-Born Decile", reverse=TRUE)) +
  theme_bw() +
  scale_color_brewer(type="div", palette="RdYlBu", direction=1)
foreign_born_par_coord
```

While some neighborhoods with a foreign population on each side of the median enjoyed better and suffered through poorer Internet infrastructure, the extremes showed a clear pattern. The top 10% of neighborhoods by foreign-born makeup were among the most mobile-dependent and among the least broadband-subscribing, while the 20% of neighborhoods with the smallest ratios of foreign-born residents subscribed to home broadband and relied on their cell phones the least for Internet access. Clearly, the most huddled masses were still yearning to browse free.

In between, things were messier. Neighborhoods in the seventh decile of foreign-born makeup subscribed to home broadband at the second highest rate of any decile. Those in the fourth subscribed the second least. Furthermore, the amount of immigrants in a neighborhood held almost no effect on the number of residents' ISP options. While there are some correlations between having extremely high or low foreign-born populations and having better or worse Internet infrastructure, in general, neighborhoods of all foreign-born makeups enjoy good and poor Internet service.

# Trying to Learn Remotely

When the Covid-19 pandemic forced society to a standstill, one of our most significant fears was its effect on children. Worries over the drops in children's learning and development of social skills loomed large in parents' thoughts, while schools muddled through teaching their students as best they could. While hard for all students, some faced more difficulties than others. Sadly, neighborhoods with the most children used some of the least reliable Internet in the city.

```{r}
data4 <- data %>% rename(
  minors = Population.under.18.years..,
  retired = Population.over.64.years..,
  Home.Broadband.Adoption = Home.Broadband.Adoption..,
  Mobile.Dependent = Mobile.Dependent.Households..Percentage.of.Households.,
  Broadband.Providers = Residential.Broadband.Choice.Average.by.NTA,
  Public.Hotspots = Public.Wi.Fi.in.NYC.Count,
  Broadband.Providers = Residential.Broadband.Choice.Average.by.NTA,
  Mobile.Broadband = Mobile.Broadband.Adoption..
)
data4$working.age <- 100 - data4$minors - data4$retired

children_vs_broadband <- ggplot(data4, aes(x=minors, y=100*Home.Broadband.Adoption)) +
  geom_point(aes(colour=Borough)) +
  geom_smooth(method='glm', size=.5, se=FALSE, color='black') +
  geom_text_repel(label=data4$Neighborhood.Name, max.overlaps=12, size=2) +
  theme_bw() +
  labs(
    title="Children's Access to Home Broadband",
    x="Neighborhood Population Younger than 18 (%)",
    y="Neighborhood Home Broadband Adoption (%)"
  ) +
  xlim(0, 40)
children_vs_broadband
```

There was a clear negative correlation between youth population and home broadband adoption. Much of this connection could be attributed to relatively child-free and better-off Manhattan's high broadband adoption. Following the time-worn connection between poverty, geography, and childbirth rates, New York neighborhoods raising a greater portion of children were the same neighborhoods with the lowest adoption of wired home Internet. For children who might not have yet been old enough to have their own cell phones, this might have meant reliance on a parent's phone hotspot to attend class or do homework.

```{r}
children_vs_hotspots <- ggplot(data4, aes(x=minors, y=Public.Hotspots)) +
  geom_point(aes(colour=Borough)) +
  geom_smooth(method='glm', size=.5, se=FALSE, color='black') +
  geom_text_repel(label=data4$Neighborhood.Name, max.overlaps=12, size=2) +
  theme_bw() +
  labs(
    title="Children's Access to Public Wi-Fi",
    x="Neighborhood Population Younger than 18 (%)",
    y="Public NYC Wi-Fi Hotspots"
  ) +
  ylim(0, 155)
children_vs_hotspots
```

If a child had to attend class remotely, then perhaps they might have been able to use a public Wi-Fi network? Unfortunately, during the pandemic, city network access points were more likely to be available in neighborhoods with fewer children. Many of the most youthful neighborhoods had 0 public hotspots at all. For students to learn effectively during Zoom school, many were not able to turn to public Wi-Fi in their neighborhood.

```{r}
data5 <- data4 %>% filter(Borough != "Manhattan")
children_vs_hotspots_no_manhattan <- ggplot(data5, aes(x=minors, y=Public.Hotspots)) +
  geom_point(aes(colour=Borough)) +
  geom_smooth(method='glm', size=.5, se=FALSE, color='black') +
  geom_text_repel(label=data5$Neighborhood.Name, max.overlaps=4, size=2) +
  theme_bw() +
  labs(
    title="Children's Access to Public Wi-Fi (Outer Boroughs)",
    x="Neighborhood Population Younger than 18 (%)",
    y="Public NYC Wi-Fi Hotspots"
  ) +
  ylim(0,60)
children_vs_hotspots_no_manhattan
```

Even excluding the high number of public hotspots in the city's Manhattan central business districts, public network connections were still more likely to be in neighborhoods with a smaller portion of children. The result was a city Internet infrastructure ill-equipped to accommodate students' transition to virtual learning.


Income and Internet Trends:
```{r, echo=FALSE}
current <- select(data, c('Borough',
                          'Per.Capita.Income..in...',
                          'Home.Broadband.Adoption..', 
                          'Residential.Broadband.Choice.Average.by.NTA',
                          'Mobile.Dependent.Households..Percentage.of.Households.' ,
                          'Blocks.with.ISP.Reported.Speed.Availability.of...900.Mbps.in..',
                          'Public.Wi.Fi.in.NYC.Count', 
                          'Mobile.Broadband.Adoption..'))

current$Per.Capita.Income..in... <-
  current$Per.Capita.Income..in.../1000

current$Mobile.Dependent.Households..Percentage.of.Households. <-
  current$Mobile.Dependent.Households..Percentage.of.Households. * 100

current$Mobile.Broadband.Adoption.. <-
  current$Mobile.Broadband.Adoption.. * 100

current$Blocks.with.ISP.Reported.Speed.Availability.of...900.Mbps.in.. <-
  current$Blocks.with.ISP.Reported.Speed.Availability.of...900.Mbps.in.. * 100

current$Home.Broadband.Adoption.. <-
  current$Home.Broadband.Adoption.. * 100


plot1 <- ggplot(current, aes(x=Per.Capita.Income..in..., y=Home.Broadband.Adoption..)) +
  geom_point(aes(color=Borough), alpha=.6, size=1) +
  geom_smooth(method='glm', size=.5, se=FALSE, color='black') +
  xlab('Income (in $1000s)') +
  ylab('% Internet Adoption') +
  theme(text = element_text(size = 10)) +
  theme_bw()
  
plot2 <- ggplot(current, aes(x=Per.Capita.Income..in..., y=Residential.Broadband.Choice.Average.by.NTA)) +
  geom_point(aes(color=Borough), alpha=.6, size=1) +
  geom_smooth(method='glm', size=.5, se=FALSE, color='black') +
  xlab('Income (in $1000s)') +
  ylab('# of Carriers') +
  theme(text = element_text(size = 10)) +
  theme_bw()

plot3 <- ggplot(current, aes(x=Per.Capita.Income..in..., y=Mobile.Dependent.Households..Percentage.of.Households.)) +
  geom_point(aes(color=Borough), alpha=.6, size=1) +
  geom_smooth(method='glm', size=.5, se=FALSE, color='black') +
  xlab('Income (in $1000s)') +
  ylab('% Mobile Dependent') +
  theme(text = element_text(size = 10)) +
  theme_bw()

plot4 <- ggplot(current, aes(x=Per.Capita.Income..in..., y=Blocks.with.ISP.Reported.Speed.Availability.of...900.Mbps.in..)) +
  geom_point(aes(color=Borough), alpha=.6, size=1) +
  geom_smooth(method='glm', size=.5, se=FALSE, color='black')  +
  xlab('Income (in $1000s)') +
  ylab('% High Speed Internet') +
  theme(text = element_text(size = 10)) +
  theme_bw()

plot5 <- ggplot(current, aes(x=Per.Capita.Income..in..., y=Public.Wi.Fi.in.NYC.Count)) +
  geom_point(aes(color=Borough), alpha=.6, size=1) +
  geom_smooth(method='glm', size=.5, se=FALSE, color='black') +
  xlab('Income (in $1000s)') +
  ylab('Public WiFi Count') +
  theme(text = element_text(size = 10)) +
  theme_bw()

plot6 <- ggplot(current, aes(x=Per.Capita.Income..in..., y=Mobile.Broadband.Adoption..)) +
  geom_point(aes(color=Borough), alpha=.6, size=1) +
  geom_smooth(method='glm', size=.5, se=FALSE, color='black') +
  xlab('Income (in $1000s)') +
  ylab('% Mobile Internt Adoption') +
  theme(text = element_text(size = 10)) +
  theme_bw()

figure <- ggarrange(plot1, plot2, plot3, plot4, plot5, plot6, ncol=3, nrow=2, common.legend = TRUE, legend="bottom")

annotate_figure(figure,
                top = text_grob("Income and Internet Trends", size=20, face = "bold"))
```

In the above graphs we can see a comparison of income levels (by borough) and several internet availability measures. The black line on each graph represents the trend between income and the respective measure of internet. The colors represent the boroughs and each dot is a neighborhood within the borough. 

As we can see from the data, Manhattan seems to be the borough with the highest income. The green dots tend to be further right and away from the other boroughs demonstrating that trend. The rest of the boroughs are relatively close together regarding their income. We see this in looking at the clusters in each graph.

Now, looking at internet trends with regards to internet availability, there are some significant patterns to be aware of. First of all, when looking at the black trend line we can see that they are all positive trends except the top right graph representing mobile dependent households. For the other five comparisons, this signifies that higher incomes are correlated with better internet. Whether it be more public Wi-Fi hot-spots, faster internet, more selection with internet providers, or higher mobile and household internet adoption. As for the top right graph showing the negative trend. This is because mobile dependency is the inverse of better internet coverage. If more households are mobile dpendent it means they do not have internet in their and therefore a lower internet adoption rate. 

Looking at these trends we can see that internet adoption is positively correlated to income level. As the higher the income the more available internet is in each neighborhood.

Significant Trend in The Bronx:
```{r, echo=FALSE}
current = select(data, c('Borough', 'Families.below.poverty.level..', 'Blocks.with.ISP.Reported.Speed.Availability.of...900.Mbps.in..'))

current$Blocks.with.ISP.Reported.Speed.Availability.of...900.Mbps.in.. <-
  current$Blocks.with.ISP.Reported.Speed.Availability.of...900.Mbps.in.. * 100
ggplot(current, aes(x=Families.below.poverty.level.., y=Blocks.with.ISP.Reported.Speed.Availability.of...900.Mbps.in.., col=Borough)) +
  geom_point(size=2) +
  xlab('Families Below Poverty Line') +
  ylab('% Internet availability > 900 Mbps') +
  ggtitle('Poverty Level and Fast Speed Internet') +
  theme_bw()
```
This graph represents families below the poverty line in each neighborhood compared to high speed internet availability. According to the US Census, the poverty line in the United States is annual income below $33,000. As for the y-axis variable, this represents the percentage of blocks in a neighborhood that have access to internet speeds of 900 Mbps or higher. This high speed internet is crucial for businesses operating several internet-dependent pieces of technology at once as well as any household with a large number of connected devices. 

As we can see from the chart The Bronx seems to be isolated from the other boroughs. No neighborhood in The Bronx has complete access to internet speeds greater than 900 Mbps. Also, we see neighborhoods int he Bronx being further right on the graph than neighborhoods in other boroughs. This signifies more families below the poverty line compared to the other boroughs. These two factors again reinforce the idea that economic status and internet adoption are correlated.


```{r}
data = read.csv(file='data/clean/combined_dataset_no_weird.csv')

data_snap <- select(data, c('Borough',
                          'Neighborhood.Name',
                          'Total.Population',
                          'Total.Number.of.Households',
                          'Per.Capita.Income..in...',
                          'Home.Broadband.Adoption..', 
                          'Residential.Broadband.Choice.Average.by.NTA',
                          'Mobile.Dependent.Households..Percentage.of.Households.' ,
                          'Public.Wi.Fi.in.NYC.Count', 
                          'Mobile.Broadband.Adoption..',
                          'English.only.language..',
                          'Language.other.than.English..',
                          'Native.born.in.US..',
                          'Foreign.born..',
                          'Completed.Bachelors.degree.or.higher..',
                          'Median.House.gross.rent..in...',
                          'Median.House.value..in...', 
                          'Owner.occupied.Housing.unit..',
                          'Total.occupied.Housing.unit..',
                          'Renter.occupied.Housing.unit..',
                          'Work.from.Home...Capable',
                          'Work.from.Home...Incapable',
                          'plurality_racial_group'
                          ))
data_snap$Per.Capita.Income..in... <- data_snap$Per.Capita.Income..in.../1000
data_snap$Mobile.Dependent.Households..Percentage.of.Households. <- data_snap$Mobile.Dependent.Households..Percentage.of.Households. * 100
data_snap$Mobile.Broadband.Adoption.. <- data_snap$Mobile.Broadband.Adoption.. * 100
data_snap$Home.Broadband.Adoption.. <- data_snap$Home.Broadband.Adoption.. * 100

```

## Do the people in NYC who we are living in rented stays have access to better internet facilities?

```{r}

# Internet Access Ratings around Rented Stays
iar <- ggplot(data, aes(fill=Borough, x=Residential.Broadband.Choice.Average.by.NTA, y=Renter.occupied.Housing.unit..)) +
    geom_line(aes(color=Borough), alpha=.5, size =8) +
    ggtitle("Internet Access Ratings around Rented Stays") +
    xlab("Region Broadband Rating by NTA") +
    ylab("Rented Housing Units in %") +
    theme_economist() + 
    scale_color_economist() +
    theme(axis.title = element_text(size = 5),
    axis.text = element_text(size = 5))

iar
```

Above plot analyses the broadband effectiveness in and around rented stays. Students, employed or unemployed people need to ensure the rent they pay give them the best facilities that's available in the region. As most of the people are immigrants, their internet connectivity concern may have a role in determining their quality of life as they may choose to live away from home.
We can observe the Brooklyn borough has wide variance in the internet connectivity spread and broadband rating in neighborhoods, however Manhattan on the other end has pretty much better access and connectivity due to more activity and internet dependent businesses that prevail in the area.The Staten island has an average NTA rating of 1.8 which is less than average of other boroughs like the Queens and Manhattan but far better than the Bronx area.


## How does the NYC household racial ethnicity distribution look like ?

```{r}

# Borough wise Household Distribution
bhd <- ggplot(data, aes(x=Borough, y=Total.Number.of.Households)) +
   geom_boxplot() + facet_wrap(~ plurality_racial_group) +
   theme_economist() + 
   scale_color_economist() +
   theme(axis.title = element_text(size = 7),
        axis.text = element_text(size = 7))

bhd
```
```{r}
# Enhanced Mosaic Plot of Racial groups residing in different boroughs
cex.axis = 0.5
adv_mos <- mosaicplot(data$Borough ~ data$plurality_racial_group, 
           main = "Mosaic Distribution of Racial Groups",
           sub = "",
           ylab = "Plurality Racial Groups",
           xlab = "Boroughs",
           las = 1,
           border = "chocolate",
           cex.axis = 0.5,
           shade = TRUE)

adv_mos
```

The neighborhood Internet infrastructure typically utilized across different racial groups vary significantly.
Double clicking on it, neighborhoods in a smaller borough with highly populated individuals get better access to infrastructure facilities, the above does showcase racial disparity . Couple of boroughs including Manhattan has fewer Hispanic groups compared to Black or African American neighborhoods, and the highest fraction of households in White groups depend and prominently access better Internet facilities than any other racial-plurality neighborhoods in NYC.

The mosaic plot also clearly demonstrates the spread and distribution in a very visual and effective manner to identify and learn the ethnicity distribution across neighborhoods affecting their social lifestyle.

## Are we democratizing Internet Access to Everyone
```{r}
# Democratized Internet Access Spread
dias <- ggplot(data, aes(Home.Broadband.Adoption..,Total.Number.of.Households)) +
  geom_point(aes(color = plurality_racial_group)) +
  theme_grey(10) +
  facet_grid(~Borough) +
  theme(strip.text = element_text(size = 10, color = "blue"),legend.position = "bottom") +
  xlab("Broadband Adoption %") +
  ylab("Number of Households in various neighborhoods") +
  ggtitle("Check for Democratized Internet Access Spread") +
  theme_wsj(base_size = 7,) +
  scale_colour_wsj("colors6") +
  theme(axis.title=element_text(size=7))

dias
```

This graph represents families in each neighborhood belonging to various ethnicity compared to high speed internet availability access . High speed access to rich and low speed to others isn't democratizing internet access to everyone, isn't it ? High speed internet is paramount for any household be  it any ethnicity.
Lot of Hispanic Latino individual neighborhoods have less internet adoption rate in the Bronx and other high visible boroughs like Manhattan.
Are we taking the steps in the right direction to democratize internet access ? We continue to identify the patterns and disparities.

## Will the rent you pay guarantee effective high internet ?

```{r}
# Home Internet Users Lifestyle
wsj_plot <- ggplot(data, aes(fill=Borough, y=Home.Broadband.Adoption.., x=Median.House.gross.rent..in...)) +
    geom_line(aes(color=Borough), alpha=.5, size =2) +
    ggtitle("Home Internet Users Lifestyle") +
    ylab("Percentage of Home Broadband Users (%) ") +
    xlab("House Gross Rent (in $)") +
    theme_wsj(base_size = 10,) +
    scale_colour_wsj("colors6") +
    theme(axis.title=element_text(size=9))

wsj_plot

```

Finding a place to stay and rent in NYC is already tough, add it to the plethora of facilities one expects to pay to leverage services like internet. The above plot insight-fully showcases that be it 1000\$ or 3000\$ rent stay, people still lack internet services to a great extent. There are still 20-30% of individuals and neighborhoods that lack better facilities of high speed internet in Queens and Brooklyn boroughs. Paying more rent may guranatee a better lifestyle but doesn't guarantee one a better high speed internet availability all the time.

## How Satisfied are internet users in HCOL (High cost of living) & LCOL (low cost of living) neighborhood areas ? 
```{r}

# Broadband Ratings in HCOL & LCOL
wsj_plot2 <- ggplot(data, aes(fill=Borough, x=Residential.Broadband.Choice.Average.by.NTA, y=Median.House.value..in...)) +
    geom_point(aes(color=Borough), alpha=.5, size =7) +
    ggtitle("Broadband Ratings in HCOL & LCOL") +
    xlab("Broadband Rating by NTA ") +
    ylab("House Value (in $)") +
    theme_wsj(base_size = 9,) +
    scale_colour_wsj("colors6") +
    theme(axis.title=element_text(size=9))

wsj_plot2

```

Brooklyn demonstrates the widespread variance in the internet availability given the high valued houses they live in. On the other hand the region of The Bronx, people are averagely happy with the services provided. However, Manhattan being so much densely populated does showcase a link between better facilities for higher house values. Looking at this graph, certainly there is direct correlation on internet access in high cost of living neighbourhoods in the dense boroughs.

## Are we moving to a mobile first era or households still rely on the high speed home internet ?
```{r}

# Parallel Coordinates Plot of Variables
ppcoord_plot <- ggparcoord(data, columns = c(4,7,6,8), alphaLines = 0.5,scale = "uniminmax", splineFactor = 10, showPoints = TRUE) +
  theme(legend.position = "none",plot.title = element_text(hjust = 0.2)) +
  scale_color_manual(values = c("grey70")) +
  ggtitle("Parallel plot of different variables for Internet Households") +
  geom_vline(xintercept = 1:4, color = "black") +
  theme(axis.text.x = element_text(angle=10))

ppcoord_plot
```

The above is a rhetorical introspection on what the future holds for NYC and the neighborhoods. The total population seemingly has direct relation to increasing family size, bigger and better homes and lifestyle. This increase in households leads to fewer increase in home broadband connections as compared to the vast increase in mobile broadband users. Across all neighborhoods in NYC 3-5 years down the line, we already are and definitely moving to a mobile first generation where prevalence of mobile phone users and internet access is gonna splurge exponentially. Households will still rely on home broadband for majority of tasks but the future awaits to see the transition and change in human behavior to mobile first.

Sources:
https://www.census.gov/topics/income-poverty/poverty/guidance/poverty-measures.html#:~:text=Step%201%3A%20Determine%20the%20family's,threshold%20(below)%20is%20%2433%2C148.